Analysis and Figures for Gallego et al, 2020

### 2020-09-06 Manuscript/What.lies.ahead.reviewed.tex
minor changes, still trying to figure out wilcox weirdness and to figure out how to easily calculate fraction of time in the future different enviro conditions occur

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the colder water and higher-pH community (community 1 in Fig. 2C) showed higher eDNA indexes of diatoms \textit{Thalassiosira nordenskioeldii}, \textit{Nitzschia cf. promare}, and \textit{Ditylum brightwellii} -- among other species -- relative to a higher-temperature community (community 3 in Fig. 2C). Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliana huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}. Community 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12\degree C and pH $<$ {\color{red} 7.9  \st{8.2}} -- where community 3 is the most likely community -- to occur more than {\color{red} four-fifths \st{two-thirds}} of the time ({\color{red} 83\%  \st{68\%}}). {\color{red} RAMON -- can you see if this 68\% stat is still true, and also if you can incorporate the similar expectations that are commented-out directly below, updating their \% if necessary?} RYAN: It is a tricky sentence as we no longer estimate seasons independently. Cluster 3 is now mode prevalent on pH < 7.8 and Temperature > 12.5; and that condition appears now 83 \% of the time in our simulation.  {\color{blue} RAMON - ah, yes -- forgot about seasons.  Could we just have some analogous thing where we say what fraction of the time we expect each community, at each area, in 2017 vs 2095?.} \\

@@

Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at richness variation within phyla; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (taxon richness by year, Wilcoxon p $=$ 0.7). Diatoms, however, show consistent and significant declines in richness in that region (Wilcoxon p $< 10^{-16}$). } {\color{blue} RAMON - can you clarify here? there should be comparisons for A. 2017 vs 2095 in overall richness, for both HC and SJI.  Then B. same for Diatom richness, dinoflagellate richness, or anything else you find interesting.}

@@

>>>>

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the colder water and higher-pH community (community 1 in Fig. 2C) showed higher eDNA indexes of diatoms \textit{Thalassiosira nordenskioeldii}, \textit{Nitzschia cf. promare}, and \textit{Ditylum brightwellii} -- among other species -- relative to a higher-temperature community (community 3 in Fig. 2C). Communities 2 and 3 had relatively higher values for taxa from a mix of groups including \textit{Emiliana huxleyi}, \textit{Minutocellus polymorphus}, \textit{Micromonas pusilla}, and \textit{Bathycoccus prasinos}. Community 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, by 2095 we expect summer in the Hood Canal to be T $>$ 12\degree C and pH $<$ {7.9} -- where community 3 is the most likely community -- to occur more than {four-fifths} of the time (83\%}). {\color{blue} I couldn't find where you were doing this calculation, so I couldn't do the fractions of time we expect the other conditions. It would be good to say that we go from condition 1 being XX\% of the time in 2017 to YY\% of the time in 2095.}

\\

@@

Given such heterogeneity in projections, gains and losses tend to balance one another out when looking at overall richness variation; {\color{red} we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5); while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $=$ 0). Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$), while the changes on the San Juan Island are undetectable (p $=$ 0.8).  Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-50}$) }.  {\color{blue} I re-did some of these in the Rmd file (starting at line 1129, doing it line-by-line to convince myself what was going on. See line 1157 in particular: for diatoms, there is no change between years, but yet the wilcox p value is very low. I can't figure out what's happening there.}

@@

{\color{blue} The paragraph below is important, and links to the question I had above: where are we doing this calculation in the code? and is the below paragraph still true? Plus, a reviewer wanted a citation of the Coscinodiscus statement.}

### 2020-09-06 Scripts/Manuscript.Main.Analysis.Rmd
minor changes, still trying to figure out wilcox weirdness and to figure out how to easily calculate fraction of time in the future different enviro conditions occur

#overall change in richness
richness.by.point %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
#diatom change in richness
richness.by.phylum.by.point %>%
  filter(phylum == "Bacillariophyta") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
richness.by.phylum.by.point %>%
  filter(phylum == "Dinophyceae") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
```

>>>>

#overall change in richness
richness.by.point %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
#diatom change in richness
richness.by.phylum.by.point %>%
  filter(phylum == "Bacillariophyta") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))
richness.by.phylum.by.point %>%
  filter(phylum == "Dinophyceae") %>%
  group_by(Year, Area) %>%
  summarise(median(richness))

#we find no change in median richness for the year 2095 relative to the present in the Hood Canal (overall taxon richness by year, averaged over all environmental conditions, Wilcoxon p $=$ 0.5);

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal") %>% 
                    pull(richness))

#while higher diversity is expected in the San Juan Island in 2095 (increase in median species richness of 2.1, Wilcoxon p $=$ 0).

wilcox.test(x = richness.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island") %>% 
                    pull(richness),
            y = richness.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island") %>% 
                    pull(richness))

#Diatoms, however, show small but significant declines in richness in the Hood Canal (0.5 species, Wilcoxon p $< 10^{-16}$),

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "Hood Canal", phylum == "Bacillariophyta") %>% 
                    pull(richness))

#while the changes on the San Juan Island are undetectable (p $=$ 0.8).

wilcox.test(x = richness.by.phylum.by.point %>%
                  filter(Year == 2017, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness),
            y = richness.by.phylum.by.point %>%
                  filter(Year == 2095, Area == "San Juan Island", phylum == "Bacillariophyta") %>% 
                    pull(richness))

#Dinoflagellates see their richness increase in both regions with the future scenario(Wilcoxon p $< 10^{-50}$)
